---
title: "Practical-Machine-Learning"
author: "Sandra Cuffee"
date: December 7, 2017
output: html_document
toc: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## **Note:**

devtools::install_github('topepo/caret/pkg/caret') , was required for parallel processing of data file.

My project can be viewed at: http://scuffee.github.io/Practical-Machine-Learning/index.html

## **Data Files**

The data files were retrieved from: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

```{r, echo = FALSE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile="pml-training.csv",method="libcurl")

download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="pml-testing.csv",method="libcurl")
```

## **The Data**

Two data frames were created in order to explore, run tests, and generate graphs pertaining to the training and test data.  The training data frame was used to create the training data set.  The testing data frame was used for the test data set. 


```{r, echo= FALSE}

training<-(read.csv("pml-training.csv",header=TRUE,sep=","))
testing<-(read.csv("pml-testing.csv",header=TRUE, sep=","))
```

## **Information About the Training and Test Data**

"Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E)."  The training data frame consisted of 19,622 observations and 160 variables.  The test data set consisted of 20 observations and 160 variables.

### Observations and Predictors
```{r,library(dplyr),library(utils),echo =FALSE}
library(dplyr);library(utils);library(caret);library(doParallel)
library(parallel);library(utils);library(ggplot2)
glimpse(training) #print variable names and types
sum(is.na(training)) #check for missing data

```
###  Cleaning the Training Data 

the training data frame was reviewed in table format using fix() and glimpse() to help determine what data would be useful for testing the training data.  The glimpse function was used to verify the predictor variables under test.  The training data set was reduced to 53 variables.  No changes were made to the test data.

```{r, echo=FALSE}

nv<-nearZeroVar(training,saveMetrics=TRUE,names = TRUE) # ID Predictors to Remove
training=training[,c(-6,-12:-17,-20,-26,-51:-59,-69:-75,-78,-79,-81,-82,-87:-92,-95,-98,-101,-125:-131,-133,-134,-139,-142:-150)]
##fix(training)
training = training[,c(-11:-25)]
##fix(training)
training = training[,c(-1:-5,-11, -12, -26,-36:-39,-43:-48,-50:-59,-72:-76,-78)]
##fix(training)
sum(is.na(training));glimpse(training)

```

## **Random Forest Analysis on Training Data Set**

Random forest tests were conducted using the caret function for accuracy and parallel processing was used to reduce the run time of the tests on the training data set. The mtry equal to 27 provided the best results for the training data set.  The accuracy was 99.68%  and the

```{r, library(caret), library(doParallel), library(parallel),echo=FALSE}
set.seed(97544)
 inTrain <- createDataPartition(training$classe,list=FALSE)
 trainingTrain<-training[inTrain,]
 testTrain<-training[-inTrain,]
 y <- trainingTrain[,54]
 x <- trainingTrain[,-54]
 cluster <- makeCluster(detectCores() - 1)
 registerDoParallel(cluster)
 fitControl <- trainControl(method = "cv",number = 5,allowParallel = TRUE)
 fit <- train(x,y, method="rf",data=training,trControl = fitControl)
 stopCluster(cluster)
 registerDoSEQ()
```

### Predictions with the Test Data
The test data consisted of 20 observations and 160 predictors.  Since the trainingSet was the basis for the test data, 53 predictors were used. The initial prediction of the test data was the raw data and the second prediction consisted of the probabilities. the raw data output shows what levels would be predicted and the probability output shows how likely the particular classe would be picked.  

```{r, echo=FALSE}
testSet<-predict(fit,testing,type="raw")
testSet1<-predict(fit,testing,type="prob")

```

### Prediction Results

Statistical analysis was completed for the training and test data sets.  The training data was resampled and cross validated for accuracy. The test data predictions was based on the training data. trainingSet represents the training data results and testSet represents the test prediction results.

```{r,library(caret), echo=FALSE}
fit
fit$resample
confusionMatrix.train(fit)
fit$finalModel
trainingSet<-(fit$finalModel)
print(trainingSet)

print(testSet);print(testSet1)

```

## **Plots**

A plot was generated for the results of each data set.

###  Training Set Diagnostics
```{r, library(ggplot2), echo=FALSE}


plot(trainingSet,pch=19,cex=0.5,col="red")
```

### Frequency of Test Predicition Levels

```{r, echo=FALSE}

plot(testSet)
```

### Probablity of Each Level

```{r, echo = FALSE}

boxplot(testSet1,col="blue")
```

## **Appendix**

### References 

gh-page setup

https://github.com/lgreski/datasciencectacontent/blob/master/markdown/pml-ghPagesSetup.md


Improving Peformance of Random Forest in caret::train()

https://github.com/lgreski/datasciencectacontent/blob/master/markdown/pml-randomForestPerformance.md


Practical Machine Learning Project

http://scuffee.github.io/Practical-Machine-Learning/index.html


Weight Lifting Exercises Dataset

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013

. 
