---
title: "Practical-Machine-Learning"
author: "Sandra Cuffee"
date: "December 7, 2017"
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 devtools::install_github('topepo/caret/pkg/caret')

## **Summary**


## **Data Files**

```{r, echo = FALSE}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile="pml-training.csv",method="libcurl")

download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile="pml-testing.csv",method="libcurl")
```

## **Data Frames**

Two data frames were created in order to explore, run tests, and generate graphs pertaining to the data sets.  The training data frame was used to create the training and vlaidation dat test sets.  The testing data frame was use st the test set in orde to perform predictions.r 


```{r, echo= FALSE}

training<-(read.csv("pml-training.csv",header=TRUE,sep=","))
testing<-(read.csv("pml-testing.csv",header=TRUE, sep=","))
```

## **Information About the Training and Test Data**

"Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E)."

### Number of Observations,Variables, and NAs
```{r,library(dplyr),library(utils),echo =FALSE}
library(dplyr);library(utils);library(caret);library(doParallel)
library(parallel);library(utils);library(ggplot2)
dim(training);dim(testing) #A count of observations and variables
glimpse(training) #print variable names and types
sum(is.na(training));sum(is.na(testing))#check for missing data

```
###  Cleaning the Training and Validation Data 

Both data frames were reviewed in table format using the fix function to help determine what data would be useful for testing the training and validation data.  The glimpse function was used to verify the predictor variables under test.

```{r, echo=FALSE}
nv<-nearZeroVar(training,saveMetrics=TRUE,names = TRUE) # ID Predictors to Remove
training=training[,c(-6,-12:-17,-20,-26,-51:-59,-69:-75,-78,-79,-81,-82,-87:-92,-95,-98,-101,-125:-131,-133,-134,-139,-142:-150)]
##fix(training)
training = training[,c(-11:-25)]
##fix(training)
training = training[,c(-1:-5,-11, -12, -26,-36:-39,-43:-48,-50:-59,-72:-76,-78)]
##fix(training)
sum(is.na(training));glimpse(training);dim(training)


glimpse(testing);dim(testing);sum(is.na(testing))

nv2<-nearZeroVar(testing,saveMetrics=TRUE,names = TRUE) # ID Predictors to Remove
#fix(testing) 
##testing=testing[,c(-1:-6,-12:-36,-50:-59,-69:-83,-87:-101,-103:-112,-125:-139,-141:-150)]
sum(is.na(testing));glimpse(testing);dim(testing) 

```

## **Random Forest Test on Training Set**

Random forest tests using the caret function and parallel processing to reduce the run time were peformed on the training and validation test sets.

```{r, library(caret), library(doParallel), library(parallrl),echo=FALSE}
set.seed(95014)

 inTraining <- createDataPartition(training$classe, p = .75, list=FALSE)
 trainingTrain <- training[inTraining,]
 testingTrain <- training[-inTraining,]
 y <- trainingTrain[,54]
 x <- trainingTrain[,-54]
 cluster <- makeCluster(detectCores() - 1)
 registerDoParallel(cluster)
 fitControl <- trainControl(method = "cv",number = 5,allowParallel = TRUE)
 fit <- train(x,y, method="rf",data=trainingTrain,trControl = fitControl)
 stopCluster(cluster)
 registerDoSEQ()
 fit
```

###  Random Forest Test on Validation Set

```{r, echo = FALSE}

set.seed(95014)

 inTraining <- createDataPartition(training$classe, p = .75, list=FALSE)
 trainingTrain <- training[inTraining,]
 testingTrain <- training[-inTraining,]
 y <- testingTrain[,54]
 x <- testingTrain[,-54]
 cluster <- makeCluster(detectCores() - 1)
 registerDoParallel(cluster)
 fitControl2 <- trainControl(method = "cv",number = 5,allowParallel = TRUE)
 fit2 <- train(x,y, method="rf",data=training,trControl = fitControl2)
 stopCluster(cluster)
 registerDoSEQ()
 fit2
```

### Random Forest Test on Test Set

The caret predict function was used on the test set.

```{r, echo=FALSE}
testSet<-predict(fit,testing)

```

 ## **Fit Models**

Statistical analysis was completed for both the training and validation sets.

```{r,library(craret), echo=FALSE}
fit
fit$resample
confusionMatrix.train(fit)
fit$finalModel
trainingFit<-(fit$finalModel)
print(trainingFit)

fit2
fit2$resample
confusionMatrix(fit2)
fit2$finalModel
validationFit<-fit2$finalModel
print(validationFit)

testSet

```

## **Plots**

The data plots for the training, validation, and test sets have been provided below.
```{r, library(ggplot2), echo=FALSE}

par(mfrow=c(1,3))
plot(fit);plot(fit2);plot(testSet)
```

## **Appendix**

### References


